---
title: "SWDIS raw analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(choroplethr)
library(choroplethrMaps)
```

# Cleaning up the water systems

Get an overview of the variables.

```{r}
read_csv("data/swdis/WATER_SYSTEM.csv", n_max=10000) %>%
  summary()
```

```{r}
water_system <- read_csv("data/swdis/WATER_SYSTEM.csv", 
                         col_types=cols_only(PWSID=col_character(),
                                             POPULATION_SERVED_COUNT=col_integer(),
                                             ORG_NAME=col_character(),
                                             PWS_TYPE_CODE=col_factor(),
                                             ZIP_CODE=col_character()
                         ))
summary(water_system)

head(water_system)
```

Cleanup some of the data.

```{r}
clean_water_system <- water_system %>%
  na.omit() %>%
  filter(POPULATION_SERVED_COUNT<1000000, POPULATION_SERVED_COUNT>500, PWS_TYPE_CODE=="CWS")
summary(clean_water_system)
```

There are still population sized outliers.

Let's look at how many people are served.

```{r}
clean_water_system %>% 
  ggplot(aes(x=POPULATION_SERVED_COUNT)) +
  geom_histogram(bins=100)
```

Let's trim off the outliers
```{r}
clean_water_system %>% 
  filter(POPULATION_SERVED_COUNT<10000) %>%
  ggplot(aes(x=POPULATION_SERVED_COUNT)) +
  geom_histogram(bins=100)
```

# Cleanup the violations

```{r}
read_csv("data/swdis/VIOLATION.csv", n_max=10000) %>% head()
```

We want to look at:
* CONTAMINANT_CODE
* PWSID
* COMPL_PER_BEGIN_DATE
* COMPL_PER_END_DATE

```{r}
violations <- read_csv("data/swdis/VIOLATION.csv",
                       col_types = cols_only(
  CONTAMINANT_CODE=col_character(),
  PWSID=col_character(),
  COMPL_PER_BEGIN_DATE=col_date(format="%d-%b-%y"),
  COMPL_PER_END_DATE=col_date(format="%d-%b-%y")
))
head(violations)
```

Get the contaminant codes.

```{r}
contaminant_codes <- read_csv("data/swdis/contaminant-codes.csv") %>%
  rename(CONTAMINANT_CODE=CODE, CONTAMINANT_NAME=NAME, CONTAMINANT_TYPE_CODE=TYPE_CODE)
head(contaminant_codes)

contaminant_group_codes <- read_csv("data/swdis/contaminant-group-codes.csv") %>%
  rename(CONTAMINANT_CODE=`Analyte Code`, CONTAMINANT_NAME=`Analyte Name`, CONTAMINANT_GROUP=`Group Name`, CONTAMINANT_GROUP_CODE=`Group Code`)
contaminant_group_codes %>% head()
```

Join this with our cleaned up PWS ids.

```{r}
clean_violations <- violations %>%
  inner_join(clean_water_system,  by="PWSID") %>%
  inner_join(contaminant_group_codes, by="CONTAMINANT_CODE")
nrow(clean_violations)
head(clean_violations)
```

Let's do some count of violation types.

```{r}
tallied_violations <- clean_violations %>%
  group_by(CONTAMINANT_NAME) %>%
  summarize(n=n()) %>%
  arrange(desc(n))

tallied_violations
```

```{r}
clean_violations %>%
  group_by(CONTAMINANT_GROUP) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(n=20)
```

Get FIPS region codes.

```{r}
data(zip_codes)
data(counties)

violations_by_region <- clean_violations %>%
  inner_join(zip_codes, by=c("ZIP_CODE"="zip")) %>%
  rename(region=fips) %>%
  group_by(region) %>%
  summarize(count_violations=n())
head(violations_by_region)
```

```{r choropleth_population, hold=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
water_system_fips <- clean_water_system %>%
  inner_join(zip_codes, by=c("ZIP_CODE"="zip")) %>%
  rename(region=fips) %>%
  group_by(region) %>%
  summarize(total_pop=sum(POPULATION_SERVED_COUNT))

county_choropleth(water_system_fips %>% rename(value=total_pop), state_zoom=c("massachusetts"))
```

```{r choropleth_population, hold=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
county_choropleth(violations_by_region %>% rename(value=total_pop), state_zoom=c("massachusetts"))
```

Why are there no violations listed for Boston itself?


```{r choropleth_population, hold=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
county_choropleth(violations_by_region %>% rename(value=count_violations), state_zoom=c("massachusetts"))
```

Why is suffolk county missing? Let's look at the boston zip code.

```{r}
clean_water_system %>%
  filter(ZIP_CODE=="02155" | ZIP_CODE=="02122")

water_system %>%
  filter(ZIP_CODE=="02155" | ZIP_CODE=="02122")

water_system %>%
  filter(grepl("BOSTON",ORG_NAME))
```

sdfsdf